---
title: "AE 15: Risk of heart disease"
subtitle: "Odds + probabilities"
date: "`r Sys.Date()`"
output: 
  html_document: 
    fig_height: 3
    fig_width: 6
  theme: readable
---


```{r setup, echo = F}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      fig.align = "center")
```


```{r load-packages, echo = F}
library(tidyverse)
library(knitr)
library(broom)
```

## Announcements

### Tea with a TA!

Hang out with the TAs from STA 210! This is a casual conversation and a fun opportunity to meet the members of the STA 210 teaching team. The only rule is these can't turn into office hours! 

Tea with a TA counts as a statistics experience.

- **Betsy Bersson**, Wed, Oct 7, 9:30 - 10:30 am
  - [Click here](https://forms.gle/naEn88fnJDEZhdVL7) to sign up. Zoom details will be emailed before the event.

## Other announcements

- **Team Feedback #1** due **Wed, Oct 7 at 11:59p**. You will receive an email from Teammates with a link to the feedback from. Email Professor Tackett if you did not receive the email (check your spam folder first!). Team feedback is part of the participation grade. 

- **Project proposal** due Fri, Oct 9 at 11:59p.

## Questions from the video? 

# High risk of heart disease

This dataset is from an ongoing cardiovascular study on residents of the town of Framingham, Massachusetts. We want to use the total cholesterol to predict if a randomly selected adult has a high risk of having coronary heart disease in the next 10 years. 

 
- `high_risk`: 
  - 1: High risk of having heart disease in next 10 years
  - 0: Not high risk of having heart disease in next 10 years

- `totChol`: total cholesterol (mg/dL)

```{r}
heart <- read_csv("data/framingham.csv") %>%
  select(totChol, TenYearCHD) %>%
  filter(!is.na(totChol)) %>% #remove observations with missing cholesterol
  mutate(high_risk = as.factor(TenYearCHD))
```

## Calculating probablity and odds

1. What is the probability a randomly selected person in the study does **<u>not</u>** have a high risk of heart disease? 

2. What are the **odds** a randomly selected person in the study does **<u>not</u>** have a high risk of heart disease? 

## Logistic regression model

1. Fit the appropriate model to predict whether a person has a high risk of heart disease given their total cholesterol. 


2. Based on the model, if a randomly selected person has a total cholesterol of 225, 
  
    - what are the log-odds they have a  high risk of heart disease? 
    - what are the **odds** they have a high risk of heart disease?   
    - what is the probability they have a high risk of heart disease? *Calculate the answer "manually" using the odds or log-odds.*
  
  
3. Based on the model, if a randomly selected person has a total cholesterol of 226, 
  
     - what are the log-odds they have a high risk of heart disease? 
    - what are the **odds** they have a high risk of heart disease?
    - what is the probability they have a high risk of heart disease? 

## Comparing log-odds and odds

1. Based on your answers in the previous section, how do the log-odds change when going from a person with total cholesterol of 225 mg/dL to a person with total cholesterol of 226 mg/dL? 

2. How would you interpret the coefficient of `totlChol` in terms of the **log-odds**?

3. What would be the interpretation of `totChol` in terms of the **odds**?
