---
title: "AE 19: Response to Leukemia Treatment"
subtitle: "Prediction"
date: "`r Sys.Date()`"
output: 
  html_document: 
    fig_height: 3
    fig_width: 6
  theme: readable
---


```{r setup, echo = F}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      fig.align = "center", 
                      echo = F)
```


```{r load-packages, echo = F}
library(tidyverse)
library(knitr)
library(broom)
library(yardstick)
```

## Announcements

### Tea with a TA!

Hang out with the TAs from STA 210! This is a casual conversation and a fun opportunity to meet the members of the STA 210 teaching team. The only rule is these can't turn into office hours! 

Tea with a TA counts as a statistics experience.

- [**Meredith Brown**](https://meredithb3.github.io/), Tue, Oct 27 at 7p
  - [Click here](https://forms.gle/fx3rUyhSYTZ9EmnE7) to sign up. Zoom details will be emailed before the event.

### This is Statistics Fall Data Challenge

[Click here](https://thisisstatistics.org/falldatachallenge/) for details on the Get out the Vote! Fall Data Challenge by the American Statistical Association (ASA). Submissions are due **November 11**.

### Big data and public policy event on Oct 26


### Final project - Draft due Oct 28

- Write the draft in the `written-report.Rmd` file in the project repo. 
- Draft should include 
   - exploratory data analysis
   - initial model selection  (main effects + interactions)
   - initial interpretations / conclusions from model
   
## Quiz 02
- Question 1.5
- Question 2
- Question 3 - D and E



## Response to Leukemia treatment

Today's data  is from a study where 51 untreated adult patients with acute myeloblastic leukemia who were given a course of treatment, and they were assessed as to their response to the treatment.^[The data set is from the Stat2Data R package] 

The goal of today's analysis is to use pre-treatment to predict how likely it is a patient will respond to the treatment. 

We will use the following variables: 

- `Age`: Age at diagnosis (in years)
- `Smear`: Differential percentage of blasts
- `Infil`: Percentage of absolute marrow leukemia infiltrate
- `Index`: 	Percentage labeling index of the bone marrow leukemia cells
- `Blasts`: Absolute number of blasts, in thousands
- `Temp`: Highest temperature of the patient prior to treatment, in degrees Fahrenheit
- `Resp`: 1 = responded to treatment or 0 = failed to respond


```{r, message = F}
leukemia <- read_csv("data/leukemia.csv")
```

1. Fit a model using `Age`, `Index` and `Temp` to predict the odds a patient will respond to the treatment. 

2. Use the augment function to obtain the predicted probability each patient in the data set will respond to the treatment.

3. Construct a confusion matrix using 0.5 as the decision-making threshold.

4. Use your confusion matrix from the previous exercise to calculate the following: 

- Sensitivity
- Specificity
- Misclassification rate

5. Create the ROC curve for this model. 

6. What is the area under the curve? Is this model a good fit for the data?

7. Suppose a doctor wants to use your model to determine if she should recommend this particular treatment for her patients with leukemia. Based on the data from the ROC curve, what decision-making threshold do you recommend the doctor use to select patients for this treatment? What factors did you consider to determine the threshold? 

