---
title: "AE 21: Sesame Street"
subtitle: "Multinomial logistic regression"
date: "`r Sys.Date()`"
output: 
  html_document: 
    fig_height: 3
    fig_width: 6
  theme: readable
---


```{r setup, echo = F}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      fig.align = "center", 
                      echo = F)
```


```{r load-packages, echo = F}
library(tidyverse)
library(knitr)
library(broom)
library(yardstick)
```

## Announcements

### Tea with a TA!

Hang out with the TAs from STA 210! This is a casual conversation and a fun opportunity to meet the members of the STA 210 teaching team. The only rule is these can't turn into office hours! 

Tea with a TA counts as a statistics experience.

- [**Meredith Brown**](https://meredithb3.github.io/), Tue, Oct 27 at 7p
  - [Click here](https://forms.gle/fx3rUyhSYTZ9EmnE7) to sign up. Zoom details will be emailed before the event.
  
- [**Cody Coombs**](www.linkedin.com/in/cody-coombs-3b8034158), Thu, Nov 5, 4p - 5p

  - [Click here](https://forms.gle/qCm63S4jUUXdnmFJ9) to sign up. Zoom details will be emailed before the event. 

### Submit your questions about statistics and the US election

What questions do you have about statistics and the US election? [Click here](https://forms.gle/Fb3rp56UcmCTKVA1A) to submit your questions **by Friday, Oct 30.** We will discuss some of the questions during the Nov 2 live lecture. 

And...if you're eligible, VOTE!

### Other events 

- Big data and public policy event TODAY 8-9 PM (Zoom info in Sakai)
- [Datathon](https://dukeml.org/datathon/) Oct 31  - Nov 1

### Stats in Spring 2021

- [STA 240: Probability for Statistical Inference, Modeling, and Data Analysis](https://stat.duke.edu/courses/probability-statistical-inference-modeling-and-data-analysis).
  - Pre-req: Calc II or Mulitvariable calc
  - Can't take it if you've already taken STA 230
  
- [STA 323: Statistical computing](https://stat.duke.edu/courses/statistical-computing)
  - [Click here](http://www2.stat.duke.edu/courses/Summer20/sta323.001-1/) 
  - Pre-req: STA 210 and STA 230 or 240

- [STA 360: Bayesian and Modern Statistics](https://stat.duke.edu/courses/bayesian-inference-and-modern-statistical-methods-0) 
  - Pre-req: STA 210, STA 230 or 240, CS 101 (or equivalent), multivariable calc
  - Co-req: Linear algebra

### Project draft - due Oct 28

- Write the draft in the `written-report.Rmd` file in the project repo. 
- Draft should include 
   - exploratory data analysis
   - initial model selection  (main effects + interactions)
   - initial interpretations / conclusions from model
   
## Sesame Street

Today's data comes from [an experiment](https://files.eric.ed.gov/fulltext/ED122799.pdf) by the Educational Testing Service to test the effectiveness of the children's program *Sesame Street*. *Sesame Street* is an educational program designed to teach young children basic educational skills such as counting and the alphabet

As part of the experiment, children were assigned to one of two groups: those who were encouraged to watch the program and those who were not.

The show is only effective if children watch it, so we want to understand what effect the encouragement had on the frequency children watched the program.

**Response:**

- `viewcat`
 + 1: rarely watched show
 + 2: once or twice a week
 + 3: three to five times a week
 + 4: watched show on average more than five times a week
 
**Predictors:**

- `age`: child's age in months
- `prenumb`: score on numbers pretest (0 to 54)
- `prelet`: score on letters pretest (0 to 58)
- `viewenc`: 1: encouraged to watch, 0: not encouraged
- `site:`
  + 1: three to five year old from urban area
  + 2: four year old from suburban area
  + 3: from rural area with high socioeconomic status
  + 4: from rural area with low socioeconomic status
  + 5: from Spanish speaking home

```{r}
# read in dataset
sesame <- read_csv("https://raw.githubusercontent.com/sta210-fa19/website/master/docs/appex/data/sesame.csv")
```

```{r}
# mean-center relevant continuous variables, make categorical variables factors
sesame <- sesame %>% 
  mutate(viewcat = factor(viewcat), 
         site = factor(site), 
         prenumbCent = prenumb - mean(prenumb), 
         preletCent = prelet - mean(prelet), 
         ageCent = age - mean(age),
         viewenc = ifelse(viewenc == 1, "1", "0"))
```
<br>

### Exploratory data analysis

1. Create a plot to visualize the relationship between the response and `viewenc`, the primary variable of interest in this analysis. What do you observe from the plot?

2. Create a plot to visualize the relationship between the response and a quantitative predictor variable. What do you observe from the plot?

### Model selection

3. Fit a model with all predictors except the primary variable of interest, `viewenc`. 

4. Conduct backward model selection using the `step` function and AIC as the model selection criterion. Display the chosen model.


### Interpretation + conclusions 

5. Interpret the intercept associated with the odds of a child being in the category `viewcat == 2` versus `viewcat == 1`. 

6. Interpret the coefficient of one numeric predictor in terms of the odds of a child being in the category `viewcat == 2` versus `viewcat == 1`. Based on the confidence interval for the coefficient, is the numeric predictor a statistically significant predictor of viewership?

7. The primary objective of the experiment was to understand the effect of encouragement `viewenc` on viewership. Does encouragement have a significant effect on viewership after adjusting for other possible factors? If so, describe the effect. Otherwise, explain why not.


