---
title: "Welcome to STA 210!"
date: "`r Sys.Date()`"
output: 
  html_document: 
    fig_height: 3
    fig_width: 6
    theme: readable
---

```{r setup, echo = F}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      fig.align = "center")
```

## Class announcements

- Office hours start Monday
- Labs start on Monday

## Check in 

## Introduction


## Course overview

Click [here](01-course-overview.html) for slides.

- Watch the videos [Welcome to Regression Analysis](https://warpwire.duke.edu/w/jQ0FAA/) and [Tour of course website + Sakai](https://warpwire.duke.edu/w/_wsFAA/) for a more detailed overview. 

## Meet your classmates! (~ 7 minutes)

- Go to a breakout room
- Introduce yourself!
  - Name
  - Year
  - What is something you're looking forward to this semester (either in this class or in general)?

## Get GitHub username (~ 5 - 7 minutse)

- Go to https://github.com/, and create an account (unless you already have one). 

- After you create your account, go to [https://forms.gle/rmJGcHio4UeiwZXh8](https://forms.gle/rmJGcHio4UeiwZXh8) and enter your NetID and  GitHub username. 

Some tips from [Happy Git and GitHub for the useR](http://happygitwithr.com/github-acct.html#username-advice) to help you create a professional GitHub usernmae.

- Incorporate your actual name.
- Reuse your username from other contexts if you can, e.g. Twitter or Slack.
- Pick a username you will be comfortable revealing to your future boss.
- Be as unique as possible in as few characters as possible. Shorter is better than longer.
- Make it timeless.</small>
- Avoid words with special meaning in programming (e.g. `NA`).


## RStudio tour 

## Tour of R and RStudio 

Let's take a tour of R and RStudio. I'll do a short demo, then give you a few minutes to try out yourself. 

### Concepts introduced

- Console
- Using R as a calculator
- Environment
- Loading and viewing a data frame
- Accessing a variable in a data frame
- R functions

### Your turn!

- Go to https://vm-manage.oit.duke.edu/containers and log into the Docker container titled "RStudio - statistics application with Rmarkdown and knitr support". This is where you'll use RStudio during the semester. 

- Use R as a calculator by typing the following in the **console**. 

```{r eval = F}
3 * 5 + 10
```

```{r eval = F}
x <- 3
x + x^2
```

```{r eval = F}
x <- 1:10
x*7
```




