---
title: "Sales vs. Advertising"
author: 
date: "08.19.20"
output: 
  html_document: 
    fig_height: 3
    fig_width: 6
---

```{r setup, echo = F}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE)
```

In this analysis, we will work with the `Advertising` data. 

## Data and packages

We start with loading the packages we'll use.

```{r load-packages, message=FALSE}
library(readr)
library(tidyverse)
library(skimr)
library(broom)
```

```{r load-data, message=FALSE}
advertising <- read_csv("data/advertising.csv")
```

We will analyze the advertising and sales data for `r nrow(advertising)` markets. The variables we'll use are 

- `tv`: total spending on TV advertising (in $thousands)
- `radio`: total spending on radio advertising (in $thousands)
- `newspaper`: total spending on newspaper advertising (in $thousands)
- `sales`: total sales (in $millions)

## Analysis

We'll begin the analysis by getting quick view of the data: 

```{r glimpse-data}
glimpse(advertising)
```

Next, we can calculate summary statistics for each of the variables in the data set. 

```{r summary-stats}
# skim() is from the skimr package
advertising %>% 
  skim()
```

1. What type of advertising typically has the smallest spending?
2. What type of advertising has the largest variation in spending? 
3. Describe the shape of the distribution of `sales`. 

We are most interested in understanding how advertising spending affect sales. One way to quantify the relationship between the variables is by calculating the correlation matrix. 

```{r correlation}
advertising %>% 
  cor()
```

1. What is the correlation between `radio` and `sales`? Interpret this value. 
2. What type of advertising has the strongest linear relationship with `sales`? 

Below are visualizations of `sales` versus each explanatory variable. 

```{r}
ggplot(data = advertising, mapping = aes(x =tv,y = sales)) + 
  geom_point(alpha=0.7) +
  geom_smooth(method="lm",se=FALSE,color="blue") + 
  labs(title = "Sales vs. TV Advertising", 
       x= "TV Advertising (in $thousands)", 
       y="Sales (in $millions") #fill in the Y axis label
```

```{r}
ggplot(data = advertising, mapping = aes(x = radio, y = sales)) + 
  geom_point(alpha = 0.7) + 
  geom_smooth(method = "lm",se=FALSE,color="red") +
  labs(title = "Sales vs. TV Advertising", 
       x= "Radio Advertising (in $thousands)", 
       y="Sales (in $millions)")
```

```{r}
## Fill in the code to create the a scatterplot sales vs. TV ads.
```

Since `tv` appears to have the strongest linear relationship with `sales`, let's calculate a simple linear regression model using these two variables. 

```{r model}
ad_model <- lm(sales ~ tv, data=advertising)
ad_model
```

1. Write the model equation. 
2. Interpret the intercept in the context of the problem. 
3. Interpret the slope in the context of the problem.

*We'll talk about slope and intercept next week!*

## Acknowledgements

The advertising data is from [*Introduction of Statistical Learning*](http://faculty.marshall.usc.edu/gareth-james/ISL/ISLR%20Seventh%20Printing.pdf)