---
title: "AE 18: Response to Leukemia Treatment"
subtitle: "Inference for logistic regression"
date: "`r Sys.Date()`"
output: 
  html_document: 
    fig_height: 3
    fig_width: 6
  theme: readable
---


```{r setup, echo = F}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      fig.align = "center", 
                      echo = F)
```


```{r load-packages, echo = F}
library(tidyverse)
library(knitr)
library(broom)
```

## Announcements

### Tea with a TA!

Hang out with the TAs from STA 210! This is a casual conversation and a fun opportunity to meet the members of the STA 210 teaching team. The only rule is these can't turn into office hours! 

Tea with a TA counts as a statistics experience.

- **Emily  **, Thu, Oct 15, 11a - 12p
  - [Click here](https://forms.gle/oUcWfeG5KD1JmGFF6) to sign up. Zoom details will be emailed before the event.

### This is Statistics Fall Data Challenge

[Click here](https://thisisstatistics.org/falldatachallenge/) for details on the Get out the Vote! Fall Data Challenge by the American Statistical Association (ASA). Submissions are due **November 11**.


## Questions from the video?

## Response to Leukemia treatment

Today's data  is from a study where 51 untreated adult patients with acute myeloblastic leukemia who were given a course of treatment, and they were assessed as to their response to the treatment.^[The data set is from the Stat2Data R package] 

The goal of today's analysis is to use pre-treatment to predict how likely it is a patient will respond to the treatment. 

We will use the following variables: 

- `Age`: Age at diagnosis (in years)
- `Smear`: Differential percentage of blasts
- `Infil`: Percentage of absolute marrow leukemia infiltrate
- `Index`: 	Percentage labeling index of the bone marrow leukemia cells
- `Blasts`: Absolute number of blasts, in thousands
- `Temp`: Highest temperature of the patient prior to treatment, in degrees Fahrenheit
- `Resp`: 1 = responded to treatment or 0 = failed to respond


```{r, message = F}
leukemia <- read_csv("data/leukemia.csv")
```

### Part 1: Age

1. Fit a model using `Age` to predict whether or not a patient responded to the treatment. *You will need to make `Resp` a factor before fitting the model.*

2. Interpret the coefficient of `Age` in the context of the data.

3. What is the distribution of the test statistic associated with `Age`? 

4. What is the conclusion in the context of the data? 

### Part 2: All pre-treatment variables

5. Now let's consider all pre-treatment variables: `Age`, `Smear`, `Infil`, `Index`, `Blasts` and `Temp`. Fit a model using these six variables to predict whether a patient responded to the treatment.

6. Based on the model, which pre-treatment variables are statistically significant? What does it mean for these variables to be "statistically significant"? 


7. Let's see if a model that only includes the significant predictors is a reasonable choice for the final model. Use a drop-in-deviance test to compare a model that includes only the significant predictors to the full model from Ex 5. Based on the results of this test, which model do you choose as the final model?




