---
title: "AE 11: Price of Diamonds, Part 2"
date: "`r Sys.Date()`"
output: 
  html_document: 
    fig_height: 3
    fig_width: 6
  theme: readable
---

```{r setup, echo = F}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      fig.align = "center")
```


## Announcements

- HW 03 due Wednesday at 11:59p (available after class)
- Introducing [statistics experiences](http://sta210-fa20.netlify.app/hw/stat-experience.html)
- Upcoming events from the [StatSci Majors Union](https://sta210-fa20.netlify.app/schedule/week-06/#announcements)
    
## HW 03

## Questions from video?

## Price of Diamonds


```{r load-packages, message=FALSE}
library(tidyverse)
library(broom)
library(patchwork)
library(knitr)
```

Today's data set contains the price and characteristics for 271diamonds randomly selected from AwesomeGems.com in July 2005.^[Data set adapted from the Diamonds data set in the Stat2Data R Package.] The variables in the data set are 

- `Carat`: Size of the diamond (in carats)
- `Color`: Coded as D (most white/bright) through J
- `Clarity`: Coded as IF (internally flawless), VVS1, VVS2, VS1, VS2, SI1, SI2, or SI3 (slightly clouded)
- `Depth`:	Depth (as a percentage of diameter)
- `PricePerCt`:	Price per carat
- `TotalPrice`: Price for the diamond (in dollars)

We will use the characteristics to understand variability in the price of diamonds. 

```{r}
diamonds <- read_csv("data/diamonds.csv")
```


## Part 1: Mean-centered variables + Interactions 

### Original model 

```{r}
orig_model <- lm(TotalPrice ~ Clarity + Carat + Clarity * Carat, data = diamonds)
tidy(orig_model, conf.int = TRUE) %>%
    kable(digits = 3)
```
### Model using mean-centered `carat`

Mean-center `Carat` and refit the model from Part 2 using the mean-centered variable for carat.

```{r}
diamonds <- diamonds %>%
  mutate(carat_cent = Carat - mean(Carat))
```

```{r}
## code to refit model model
mean_cent_model <- lm(TotalPrice ~ Clarity + carat_cent + Clarity*carat_cent , data = diamonds)
tidy(mean_cent_model, conf.int = TRUE) %>%
    kable(digits = 3)
```

1. Interpret the intercept in the context of the data. 

2. Let's compare the original and mean-centered model for `ClarityIF`. 

3. Let's compare the original and mean-centered models for `ClaritySI1`.

### Testing coefficients (Zoom poll)

Let's use the mean-centered version of the model to answer the following questions: 

1. Suppose we wish to test if the mean price for diamonds with `Clarity = "SI3"` differs significantly from diamonds with `Clarity = "IF"`. We should use the inferential statistics for which term to conduct this test?

(a) Intercept
(b) ClaritySI3
(c) carat_cent
(d) ClaritySI3:carat_cent

2. Suppose we wish to test if the effect of carat for diamonds with `Clarity = "SI3"` differs significantly from diamonds with `Clarity = "IF"`. We should use the inferential statistics for which term to conduct this test?

(a) Intercept
(b) ClaritySI3
(c) carat_cent
(d) ClaritySI3:carat_cent

3. What are the degrees of freedom associated with the regression standard error, $\hat{\sigma}$, and therefore the degrees of freedom associated with the two tests mentioned above? *Note: There are 271 observations in the diamonds data set.*

(a) 269
(b) 268
(d) 256
(c) 255


4. What is the 95% confidence interval for the amount by which the slope of carat for diamonds with `Clarity = VS1` differs from the slope of carat for diamonds with `Clarity = IF`?

(a) (4659.287, 5691.504)
(b) (-1110.613,	-26.858)
(c) (6298.339, 9243.176)
(d) (283.132, 3482.653)

## Transformations

Let's take a look at the plot of residuals vs. fitted.  

```{r}
model_aug <- augment(mean_cent_model)
```

```{r}
ggplot(data = model_aug, aes(x = .fitted, y  = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "Red") +
  labs(x = "Predicted", 
      y = "Residual", 
      title = "Residuals vs. Predcited")
```


1. Based on what you learned about conditions for SLR, which condition(s) appear to be violated? *Select all that apply.*

(a) Linearity
(b) Constant variance
(c) Normality 
(d) Independence

2. Let's use a log-transformation on the total price and refit the model.

```{r}
diamonds <- diamonds %>%
  mutate(log_total_price = log(TotalPrice))
```

```{r}
log_model <- lm(log_total_price ~ Clarity + carat_cent + Clarity*carat_cent, data = diamonds)
tidy(log_model, conf.int = TRUE) %>%
  kable(digits = 3)
```

3. Interpret the intercept in terms of `log(TotalPrice)`. 

4. Interpret the intercept in terms of `TotalPrice`. 

5. Let's focus on diamonds with `Clarity = IF`.   
  - Write the model in terms of the `log(TotalPrice)`. 
  - Interpret the coefficient of `carat_cent` in terms of `log(TotalPrice)`
  - Interpret the coefficient of `carat_cent` in terms of the `TotalPrice`.

6. Let's check the plot of residuals vs. fitted for this model. 

```{r}
log_model_aug <- augment(log_model)
```

```{r}
ggplot(data = log_model_aug, aes(x = .fitted, y  = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "Red") +
  labs(x = "Predicted", 
      y = "Residual", 
      title = "Residuals vs. Predcited")
```

It looks like we fixed the constant variance issue, but linearity is still violated!


## Higher-order terms

Let's take a look at a plot of log(TotalPrice) vs. carat and the residuals vs. carat. 

```{r}
ggplot(data = diamonds, aes(x = carat_cent, y =log_total_price )) +
  geom_point() +
  labs(x= "Mean-Centered Carat", 
       y  = "Log-transformed Total Price)", 
       title = "Log-transformed total price vs. Mean-Centered Carat")
```

```{r}
ggplot(data = log_model_aug, aes(x = carat_cent, y = .resid)) +
  geom_point() + 
  geom_hline(yintercept = 0, color = "red") +
  labs(x = "Mean-Centered Carat", 
       y = "Residual", 
       title = "Residuals vs. Mean-Centered Carat")
```


1. Let's add a quadratic term to the model. 

```{r}
log_model_v2 <- lm(log_total_price ~ Clarity + carat_cent + Clarity*carat_cent + I(carat_cent^2), 
                   data = diamonds)
tidy(log_model_v2, conf.int = TRUE) %>%
  kable(digits = 3)
```

2. Write the model equation for `Clarity = IF`. 

3. Let's interpret the effect of carat on the total price for diamonds with Clarity = IF that have a carat size within 0.2 carats of the mean (~Q1 and Q3).



 
