---
title: "AE 10: Price of diamonds"
date: "`r Sys.Date()`"
output: 
  html_document: 
    fig_height: 3
    fig_width: 6
  theme: readable
---

```{r setup, echo = F}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      fig.align = "center")
```


## Announcements

- HW 03 due Sep 23 at 11:59p (available after class)
- Zoom waiting room for all class meetings going forward
- **Upcoming event:**: [ICPSR Data Fair: Data in Real Life](https://www.icpsr.umich.edu/web/pages/membership/datafair/)     
    - Sep 21 - 25
    - Free to attend but you need to register
    - Can count towards a Stats Experience (more on these Monday!)
    
## Questions from video?

## Price of Diamonds


```{r load-packages, message=FALSE}
library(tidyverse)
library(broom)
library(patchwork)
library(knitr)
```

Today's data set contains the price and characteristics for 271diamonds randomly selected from AwesomeGems.com in July 2005.^[Data set adapted from the Diamonds data set in the Stat2Data R Package.] The variables in the data set are 

- `Carat`: Size of the diamond (in carats)
- `Color`: Coded as D (most white/bright) through J
- `Clarity`: Coded as IF (internally flawless), VVS1, VVS2, VS1, VS2, SI1, SI2, or SI3 (slightly clouded)
- `Depth`:	Depth (as a percentage of diameter)
- `PricePerCt`:	Price per carat
- `TotalPrice`: Price for the diamond (in dollars)

We will use the characteristics to understand variability in the price of diamonds. 

```{r}
diamonds <- read_csv("data/diamonds.csv")
```


## Part 1: Categorical predictors

### Model with single categorical predictor

Let's fit a model using `Clarity` to predict the price.

```{r}
model1 <- lm(TotalPrice ~ Clarity, data = diamonds)
tidy(model1) %>%
  select(term, estimate) %>%
  kable(digits = 3)
```

1. What is the baseline level? 

2. What is the interpretation of `ClaritySI1`?

3. What is the expected price of a diamond with `ClarityVVS2`?

4. What is the difference in the expected price between a diamond with `ClaritySI3` and a diamond with `ClarityVVS1`? 

### Change baseline 

We can change the baseline category using the `fct_relevel` function in the forcats package.

```{r}
diamonds <- diamonds %>%
  mutate(Clarity = fct_relevel(Clarity, c("SI3", "IF", "VVS1", "VVS2", "VS1", "VS2", "SI1", "SI2"))
  )
```

Let's refit the model: 

```{r}
model1_relevel <- lm(TotalPrice ~ Clarity, data = diamonds)
tidy(model1_relevel) %>%
  kable(digits = 3)
```

5. Interpret the coefficient for `ClarityVVS1`. 

6. How does the coefficient for `ClarityVVS1` compare to your response to Exercise 4 above? Is this what you expect? Why or why not?


## Part 2: Interaction terms

Now let's fit a model using `Clarity`, `Carat`, and the interaction between the two variables. 

```{r}
model2 <- lm(TotalPrice ~ Clarity + Carat + Clarity * Carat, data = diamonds)
tidy(model2) %>%
  kable(digits = 3)
```


1. Write the model equation for a diamond with `ClaritySI3`. 

2. Interpret the coefficient of `ClarityVVS1:Carat`. 

3. Write the model equation for a diamond with `ClarityVVS1`. 

4. Describe the effect of carat on the price of a diamond with `ClarityVVS1`.


## Part 3: Mean-center variables

Mean-center `Carat` and refit the model from Part 2 using the mean-centered variable for carat.

```{r}
#write code to mean-center

```

```{r}
## code to refit model model

```


1. Interpret the coefficient for the mean-centered version of carat.

2. Interpret the intercept. 







