---
title: "AE 22: Multinomial logistic Q+A and Project"
subtitle: "Multinomial logistic regression"
date: "`r Sys.Date()`"
output: 
  html_document: 
    fig_height: 3
    fig_width: 6
  theme: readable
---


```{r setup, echo = F}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      fig.align = "center", 
                      echo = F)
```


```{r load-packages, echo = F}
library(tidyverse)
library(knitr)
library(broom)
library(yardstick)
```

## Announcements

### Tea with a TA!

Hang out with the TAs from STA 210! This is a casual conversation and a fun opportunity to meet the members of the STA 210 teaching team. The only rule is these can't turn into office hours! 

Tea with a TA counts as a statistics experience.

- [**Cody Coombs**](www.linkedin.com/in/cody-coombs-3b8034158), Thu, Nov 5, 4p - 5p

  - [Click here](https://forms.gle/qCm63S4jUUXdnmFJ9) to sign up. Zoom details will be emailed before the event. 

### Submit your questions about statistics and the US election

What questions do you have about statistics and the US election? [Click here](https://forms.gle/Fb3rp56UcmCTKVA1A) to submit your questions **by Friday, Oct 30.** We will discuss some of the questions during the Nov 2 live lecture. 

And...if you're eligible, VOTE! Find out more at https://vote.duke.edu/


### Other events 

- [Datathon](https://dukeml.org/datathon/) Oct 31  - Nov 1

### Stats in Spring 2021

- [STA 240: Probability for Statistical Inference, Modeling, and Data Analysis](https://stat.duke.edu/courses/probability-statistical-inference-modeling-and-data-analysis).
  - Pre-req: Calc II or Mulitvariable calc
  - Can't take it if you've already taken STA 230
  
- [STA 323: Statistical computing](https://stat.duke.edu/courses/statistical-computing)
  - [Click here](http://www2.stat.duke.edu/courses/Summer20/sta323.001-1/) 
  - Pre-req: STA 210 and STA 230 or 240

- [STA 360: Bayesian and Modern Statistics](https://stat.duke.edu/courses/bayesian-inference-and-modern-statistical-methods-0) 
  - Pre-req: STA 210, STA 230 or 240, CS 101 (or equivalent), multivariable calc
  - Co-req: Linear algebra

### Project draft - due Oct 28

- Write the draft in the `written-report.Rmd` file in the project repo. 
- Draft should include 
   - exploratory data analysis
   - initial model selection  (main effects + interactions)
   - initial interpretations / conclusions from model
   
## Questions about multinomial logistic regression 

## Project office hours


**There is no AE to turn in for today's class.**