---
title: "AE 20: Response to Leukemia Treatment"
subtitle: "Prediction + Conditions"
date: "`r Sys.Date()`"
output: 
  html_document: 
    fig_height: 3
    fig_width: 6
  theme: readable
---


```{r setup, echo = F}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      fig.align = "center", 
                      echo = F)
```


```{r load-packages, echo = F}
library(tidyverse)
library(knitr)
library(broom)
library(yardstick)
```

## Announcements

### Tea with a TA!

Hang out with the TAs from STA 210! This is a casual conversation and a fun opportunity to meet the members of the STA 210 teaching team. The only rule is these can't turn into office hours! 

Tea with a TA counts as a statistics experience.

- [**Meredith Brown**](https://meredithb3.github.io/), Tue, Oct 27 at 7p
  - [Click here](https://forms.gle/fx3rUyhSYTZ9EmnE7) to sign up. Zoom details will be emailed before the event.

### This is Statistics Fall Data Challenge

[Click here](https://thisisstatistics.org/falldatachallenge/) for details on the Get out the Vote! Fall Data Challenge by the American Statistical Association (ASA). Submissions are due **November 11**.

### Big data and public policy event on Oct 26

### Final project - Draft due Oct 28

- Write the draft in the `written-report.Rmd` file in the project repo. 
- Draft should include 
   - exploratory data analysis
   - initial model selection  (main effects + interactions)
   - initial interpretations / conclusions from model
   
### Quiz 03
- Available Thu, Oct 22 at 12a  - Fri, Oct 23 11:59p
- 2 hours to complete it
- Covers material weeks 07 - 09 
- No office hours Oct 22 - Oct 23. Piazza will be inactive while the quiz is out.
- No lab this week.

## Response to Leukemia treatment

Today's data  is from a study where 51 untreated adult patients with acute myeloblastic leukemia who were given a course of treatment, and they were assessed as to their response to the treatment.^[The data set is from the Stat2Data R package] 

The goal of today's analysis is to use pre-treatment to predict how likely it is a patient will respond to the treatment. 

We will use the following variables: 

- `Age`: Age at diagnosis (in years)
- `Smear`: Differential percentage of blasts
- `Infil`: Percentage of absolute marrow leukemia infiltrate
- `Index`: 	Percentage labeling index of the bone marrow leukemia cells
- `Blasts`: Absolute number of blasts, in thousands
- `Temp`: Highest temperature of the patient prior to treatment, in degrees Fahrenheit
- `Resp`: 1 = responded to treatment or 0 = failed to respond


```{r, message = F}
leukemia <- read_csv("data/leukemia.csv") %>%
  mutate(Resp = factor(Resp))
```

We want to use the `Age`, `Index` and `Temp` to predict the odds a patient will respond to the treatment. The model is below. 

```{r}
resp_model <- glm(Resp ~ Age + Index + Temp, data = leukemia, 
                  family = "binomial")

tidy(resp_model) %>%
  kable(digits = 3)
```

Ultimately, we want to use the model to classify patients into two groups: those who are likely to respond to the treatment and those are not. To do so, we'll fit an ROC curve to help us determine a decision-making threshold. 

```{r}
resp_aug <- augment(resp_model) %>%
  mutate(Resp = fct_relevel(Resp, c("1", "0")))
```

```{r}
# calculate sensitivity and specificity at each threshold
roc_curve_data <- resp_aug %>%
  roc_curve(Resp, .fitted) %>%
  mutate(pred_prob = exp(.threshold)/(1 + exp(.threshold)))

roc_curve_data %>%
  slice(1:5)
```
```{r}
autoplot(roc_curve_data)
```


1. Suppose a doctor wants to use your model to determine if she should recommend this particular treatment for her patients with leukemia. Based on the data from the ROC curve, what decision-making threshold do you recommend the doctor use to select patients for this treatment? What factors did you consider to determine the threshold? 


2. Check the model conditions - linearity, randomness, independence. *Recall you may need to install the Stat2Data package.*


