---
title: "AE 25: Missing data"
date: "`r Sys.Date()`"
output: 
  html_document: 
    fig_height: 3
    fig_width: 6
  theme: readable
  toc: true
  toc_depth: 2
---


```{r setup, echo = F}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      fig.align = "center", 
                      echo = F)
```


```{r load-packages, echo = F}
library(tidyverse)
```

## Announcements

### Assignments 

- [Written report](https://sta210-fa20.netlify.app/project/#written-report) due Tue, Nov 17
- [Video presentation](https://sta210-fa20.netlify.app/project/#presentation-slides) due Fri, Nov 20
  - [Presentation slides](https://sta210-fa20.netlify.app/project/#presentation-slides)
- [Final repo](https://sta210-fa20.netlify.app/project/#project-repository) due Fri, Nov 20
  - Includes codebook (a list of variables and their definitions) in the `data` folder
- [Watch and comment on presentations](https://sta210-fa20.netlify.app/project/#peer-feedback) Nov 20 - Nov 22

## Exercises 

## Dealing with Missing Data

We will use the `nhanes2` data set from the mice R package. This data set is small subset of the NHANES data set specifically used to demonstrate imputation methods. 

```{r echo = T}
library(tidyverse)
nhanes2 <- mice::nhanes2
```

## bmi 

1. Let's take a look at the variable `bmi` (body mass index). 
- How many observations have missing values for `bmi`? 
- Visualize the distribution of `bmi`. 
- What is the standard deviation of `bmi` for the observations that have values for `bmi`?

2. Impute the missing values of `bmi` using mean imputation. 

3. Visualize the distribution of `bmi` with the imputed values and calculate the standard deviation. How did the distribution of `bmi` change when we filled in missing values using mean imputation? 

4. What are some potential limitations of mean imputations? 

## hyp

1. Let's consider the variable `hyp` (hypertension). How many observations have missing values for `hyp`? 

2. What are two strategies you can use to impute values for `hyp`? 

3. What are the advantages and potential limitations of the strategies you proposed?

