---
title: "Data analysis + modeling"
author: "Prof. Maria Tackett"
output:
  xaringan::moon_reader:
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    css: "sta210-slides.css"
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: "%current%"
      ratio: "16:9"
---

```{r setup, include=FALSE}
# R options
options(
  htmltools.dir.version = FALSE, # for blogdown
  show.signif.stars = FALSE,     # for regression output
  warm = 1
  )
# Set dpi and height for images
library(knitr)
opts_chunk$set(fig.height = 2.65, dpi = 300) 
# ggplot2 color palette with gray
color_palette <- list(gray = "#999999", 
                      salmon = "#E69F00", 
                      lightblue = "#56B4E9", 
                      green = "#009E73", 
                      yellow = "#F0E442", 
                      darkblue = "#0072B2", 
                      red = "#D55E00", 
                      purple = "#CC79A7")

options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
	fig.align = "center",
	fig.height = 3.5,
	fig.width = 5.8333,
	message = FALSE,
	warning = FALSE
)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(readr) #read_csv() function
library(patchwork)
```

```{r data, echo  = F}
advertising <- read_csv("data/advertising.csv")
```

## Sales vs. Advertising 

- Suppose you are a data scientist on the marketing team and the company wants to improve the sales of their premiere product 

- You want to understand the relationship between .vocab[advertising budget] and .vocab[total sales]

- The goal is to advise the marketing team about how to set the advertising budget based on their target sales goals

---

## Advertising vs. Sales

```{r glimpse}
glimpse(advertising)
```

--

- .vocab[Observations]: 200 markets

--

- .vocab[Variables]:
    - `tv`: Spending on TV ads (in $thousands)
    - `radio`: Spending on radio ads (in $thousands)
    - `newspaper`: Spending on newspaper ads (in $thousands)
    - `sales`: total sales (in $millions)
   
---

## Terminology

- `sales` is the .vocab[response variable]
    - variable whose variation we want to understand / variable we wish to predict
    - also known as *outcome* or *dependent* variable

--

<br> 

- `tv`, `radio`, `newspaper` are the .vocab[predictor variables]
    - variables used to account for variation in the outcome
    - also known as *explanatory*, *independent*, or *input* variables

---

## Let's look at the data

```{r echo=F, fig.height = 2.25}
p1 <- advertising %>% 
  ggplot(mapping=aes(x=tv,y=sales)) + 
  geom_point(alpha=0.5) + 
  labs(x = "TV Ads (in $1000s)", 
       y = "Sales (in $millions)") + 
  geom_smooth(method="lm", color="blue",se=FALSE) 

p2 <- advertising %>% 
  ggplot(mapping=aes(x=radio,y=sales)) + 
  geom_point(alpha=0.5) + 
  labs(x = "Radio Ads (in $1000s)", 
       y = "Sales (in $millions)") + 
  geom_smooth(method="lm", color="red", se=FALSE) 

p3 <- advertising %>% 
  ggplot(mapping=aes(x=newspaper,y=sales)) + 
  geom_point(alpha=0.5) + 
  labs(x = "Newspaper (in $1000s)", 
       y = "Sales (in $millions)") + 
  geom_smooth(method="lm", color="purple", se=FALSE) 

p1 + p2 + p3
```

--

Each line represents model we could use to predict `sales` using `tv`, `radio`, or `newspaper`

---

## Let's look at the data

```{r echo=F, fig.height = 2.25}
p1 <- advertising %>% 
  ggplot(mapping=aes(x=tv,y=sales)) + 
  geom_point(alpha=0.5) + 
  labs(x = "TV Ads (in $1000s)", 
       y = "Sales (in $millions)") + 
  geom_smooth(method="lm", color="blue",se=FALSE) 

p2 <- advertising %>% 
  ggplot(mapping=aes(x=radio,y=sales)) + 
  geom_point(alpha=0.5) + 
  labs(x = "Radio Ads (in $1000s)", 
       y = "Sales (in $millions)") + 
  geom_smooth(method="lm", color="red", se=FALSE) 

p3 <- advertising %>% 
  ggplot(mapping=aes(x=newspaper,y=sales)) + 
  geom_point(alpha=0.5) + 
  labs(x = "Newspaper (in $1000s)", 
       y = "Sales (in $millions)") + 
  geom_smooth(method="lm", color="purple", se=FALSE) 

p1 + p2 + p3
```

.alert[
$$\text{sales} = f(\text{tv}, \text{radio}, \text{newspaper}) + \epsilon$$
]

---

## Model 

.alert[
$$\text{sales} = f(\text{tv}, \text{radio}, \text{newspaper}) + \epsilon$$
]

- **Goal**: Define $f$

--

- How do we define $f$?
  - Make an assumption about the functional form $f$
  - Use the data to fit a model based on that form 
    
---

## General form of model

.alert[
$$Y = f(\mathbf{X}) + \epsilon$$
]

- $Y$: quantitative response variable

- $\mathbf{X} = (X_1, X_2, \ldots, X_p)$: predictor variables

- $f$: fixed but unknown function

- $\epsilon$: random error term with mean 0 that is independent of $\mathbf{X}$ 
   
---

## How to define $f$

In general, 

1. Choose the functional form of $f$, i.e. <font class="vocab">choose the appropriate model given the data</font>
    - Ex: $f$ is a linear model
    $$f(\mathbf{X}) = \beta_0 + \beta_1 X_1 + \dots + \beta_p X_p + \epsilon$$

--

2. Use the data to fit (or train) the model, i.e. <font class="vocab">estimate the model parameters</font>
    - Ex: Find estimates of $\beta_0, \beta_1, \ldots, \beta_p$

---

## Why?

.alert[
$$\hat{\text{sales}} = \hat{\beta}_0 + \hat{\beta}_1 \times \text{tv} + \hat{\beta}_2 \times \text{radio} + \hat{\beta}_3 \times \text{newspaper}$$
]

--

.vocab[Prediction:]

What do we expect `sales` to be in a market where there is $100,000 spent on TV ads, $30,000 spent on radio ads, and $10,000 spent on newspaper ads?

--

.vocab[Inference:]

What is the relationship between spending on TV ads an sales after accounting for spending on radio and newspaper ads?

---

## Course Outline

.pull-left[

.vocab[Unit 1: Quantitative Response Variables]
  - Simple Linear Regression 
  - Multiple Linear Regression

<br> 

.vocab[Unit 3: Looking Ahead]
  - Weighted least squares
  - Dealing with messy data
  - Modeling in practice

]

.pull-right[

- .vocab[Unit 2: Categorical Response Variable]
  - Logistic Regression 
  - Multinomial Logistic Regression 

]


    