---
title: "Multiple linear regression"
subtitle: "Model Conditions"
author: "Prof.  Maria Tackett"
output:
  xaringan::moon_reader:
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    css: "sta210-slides.css"
    logo: img/sta199-sticker-icon.png
    lib_dir: libs/font-awesome
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: "%current%" 
      ratio: "16:9"
editor_options: 
  chunk_output_type: console
---

```{r child = "setup.Rmd"}
```

class: middle, center

## [Click for PDF of slides](09-mlr-conditions.pdf)


---


class: middle, center

## Checking model conditions

```{r echo = F}
library(tidyverse)
library(knitr)
library(broom)
library(patchwork) 
library(kableExtra)
```

```{r,echo=F}
homes <- read_csv("data/homeprices.csv")
```

---

## Conditions

Inference on the regression coefficients and predictions are reliable only when the regression assumptions are reasonably satisfied: 

1. <font class="vocab">Linearity: </font> Response variable has a linear relationship with the predictor variables in the model

2. <font class="vocab">Constant Variance: </font>The regression variance is the same for all set of predictor variables $(X_1, \ldots, X_p)$

3. <font class="vocab">Normality: </font> For a given set of predictors $(X_1, \ldots, X_p)$, the response, $Y$, follows a Normal distribution around its mean 

4. <font class="vocab">Independence: </font>All observations are independent

--
.alert[
We will use plots of the residuals to check these assumptions 
]

---

## Checking linearity 

```{r}
price_aug <- augment(price_model)
```


```{r}
ggplot(data = price_aug, aes(x = .fitted, y = .resid)) +
  geom_point() + 
  geom_hline(yintercept = 0, color = "red") +
  labs(x = "Predicted", y = "Residuals", 
       title = "Residuals vs. Predicted")
```

---

## Checking linearity

```{r}
library(GGpairs)
```


---

## Checking constant variance assumption 

- Make a plot of the residuals vs. the predicted (fitted) values

- The height of the cloud of points should be constant as you go from left to right on the plot

---

## Checking normality assumption 

- Make the following plots: 
    - Histogram of the residuals
    - Normal QQ-Plot of the residuals 

- The histogram should be approximately unimodal and symmetric. 

- The points on the Normal QQ-Plot should generally follow a straight diagonal line
    
---

## Checking independence assumption 

- In the indepednece assumption, we assume the residuals are not correlated

- If your data were collected over time, plot the residuals in time order 

- There should be no pattern in the plot.
  - A cyclical pattern indicates the residuals are correlated, a violation of the assumption. 
  
- Can generally conclude this assumption is resonably met unless there are clear violations

---

## Check linearity: Residuals vs. Predicted

```{r}
ggplot(data = loans_aug, aes(x = .fitted, y = .resid)) +
  geom_point(alpha = 0.3) + 
  geom_hline(yintercept = 0, color = "red") +
  labs(x = "Predicted", y = "Residuals", 
       title = "Residuals vs. Predicted")
```

---

## Check linearity: Residuals vs. Predictors

```{r echo = F, fig.height = 6, fig.width = 9}
p2 <- ggplot(data = loans_aug, aes(x = verified_income, y = .resid)) +
  geom_boxplot() + 
  geom_hline(yintercept = 0, color = "red") +
  labs(y = "Residuals")

p3 <- ggplot(data = loans_aug, aes(x = debt_inc_cent, y = .resid)) +
  geom_point(alpha = 0.3) + 
  geom_hline(yintercept = 0, color = "red") +
  labs(x = "Debt to Income", y = "Residuals")

p4 <- ggplot(data = loans_aug, aes(x = bankruptcy, y = .resid)) +
  geom_boxplot() + 
  geom_hline(yintercept = 0, color = "red") +
  labs(y = "Residuals")

p5 <- ggplot(data = loans_aug, aes(x = term_cent, y = .resid)) +
  geom_point(alpha = 0.3) + 
  geom_hline(yintercept = 0, color = "red") +
  labs(x = "Term", y = "Residuals")

p6 <- ggplot(data = loans_aug, aes(x = credit_util_cent, y = .resid)) +
  geom_point(alpha = 0.3) + 
  geom_hline(yintercept = 0, color = "red") +
  labs(x = "Credit Utility", y = "Residuals")

p2 + p3 + p4 + p5 + p6
```

---

## Check constant variance

```{r echo  = F}
ggplot(data = loans_aug, aes(x = .fitted, y = .resid)) +
  geom_point(alpha = 0.3) + 
  geom_hline(yintercept = 0, color = "red") +
  labs(x = "Predicted", y = "Residuals", 
       title = "Residuals vs. Predicted")
```


---

## Check Normality


```{r echo = F, fig.height = 6, fig.width = 9}
p1 <- ggplot(data = loans_aug, aes(x = .resid)) +
  geom_histogram() +
  labs(x = "Residuals", y = "", 
       title = "Distribution of Residuals")

p2 <- ggplot(data = loans_aug, aes(sample = .resid)) +
  stat_qq() + 
  stat_qq_line() + 
  labs(title = "Normal QQ-Plot of Residuals")

p1 + p2 #using patchwork package
```


---

## Checking independence

Can check residuals versus observation number if you think there is some structure / order to the dataset. Below is the code for this dataset:

.small[
```{r}
loans_aug <- loans_aug %>%
  mutate(obs_num = 1:nrow(loans_aug))
```

```{r, fig.height = 2.5}
ggplot(data = loans_aug, aes(x = obs_num, y = .resid)) + 
  geom_point(alpha = 0.3) +
  labs(x = "Observation Number", y = "Residuals", 
       title = "Residuals vs. Observation Number")
```
]

---

## Use EDA but don't solely rely on it

- Look at a scatterplot of the response variable vs. each of the predictor variables in the exploratory data analysis before calculating the regression model 

- This is a good way to check for obvious departures from linearity or constant variance

- This is **<u>not</u>** definitive, but it can give you an indication early on if you might need to use interactions, higher-order terms, or do a transformation (more on that next week)




