---
title: "Lab 06: Logistic regression"
subtitle: "due Wed, Oct 21 at 11:59p"
output: 
  tufte::tufte_html:
    css: "./sta210-labs.css"
    tufte_variant: "envisioned"
    highlight: pygments
    fig_height: 10
    toc: true
    toc_depth: 1
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning=FALSE, 
                      message=FALSE,
                      eval=FALSE)
```

The General Social Survey (GSS) has been used to measure trends in attitudes and behaviors in American society since 1972. In addition to collecting demographic information, the survey includes questions used to gauge attitudes about government spending priorities, confidence in institutions, lifestyle, and many other topics. A full description of the survey may be found [here](http://www.norc.org/Research/Projects/Pages/general-social-survey.aspx).

In today's lab, we will use logistic regression to understand the relationship between a person's political views and their attitudes towards government spending on mass transportation projects. To do so, we will use data from the 2010 GSS survey.

# Getting Started 

- Go to the sta210-fa20 organization on GitHub ([https://github.com/sta210-fa20](https://github.com/sta210-fa20)). Click on the repo with the prefix **lab-06-**. It contains the starter documents you need to complete the lab.

- Clone the repo and create a new project in your [RStudio Docker Container](https://vm-manage.oit.duke.edu). 

- Configure git by typing the following in the **console**.

```{r eval=FALSE}
library(usethis)
use_git_config(user.name="github username", user.email="your email")
```

## Password caching

If you would like your git password cached for a week for this project, type the following in the **Terminal** in RStudio:

```{bash eval=FALSE}
git config --global credential.helper 'cache --timeout 604800'
```

You will need to enter your GitHub username and password one more time after caching the password. After that you won't need to enter your credentials for 604800 seconds = 7 days.


## Packages

We will use the following packages in today's lab. Feel free to add any other packages as needed.

```{r load-packages}
library(tidyverse)
library(broom)
library(knitr)
# fill in other packages as needed
```


## Data

The data for this lab are from the 2016 General Social Survey. The original data set contains 2867 observations and 935 variables. Given the size of the dataset, we will handle it differently in our workflow than we've handled data in previous assignments. 

```{marginfigure}
[Working with large files](https://help.github.com/en/articles/working-with-large-files)
```

The size of this dataset is 34.3 MB. GitHub will not allow you to push files larger than 100 MB and will give you a warning when you push files as large as 50 MB. Though we could push the file we're working with today to GitHub, it's large enough that we'd still prefer not to. 

You have may noticed that each repo contains a file called `.gitignore`. It contains a list of the files you don't want commit or push to GitHub. If you look at the `.gitignore` file for today's lab, you will notice that `gss2016.csv` is listed at the bottom. 

- Click [**here**](https://sta210-fa20.netlify.app/data/gss2016.csv) to download `gss2016.csv`.
- Upload `gss2016.csv` into the data folder of your project. 
- Notice that `gss2016.csv` does not appear in your Git pane. This is because it is being ignored by git, since it is listed in the `.gitignore` file.

You will use the following variables in the lab: 

- `natmass`: Respondent's answer to the following prompt:

    "We are faced with many problems in this country, none of which can be solved easily or inexpensively. I’m going to name some of these problems, and for each one I’d like you to tell me whether you think we’re spending too much money on it, too little money, or about the right amount...are we spending too much, too little, or about the right amount on mass transportation?"

- `age`: Age in years.
- `sex`: Sex recorded as *male* or *female*
- `sei10`: Socioeconomic index from 0 to 100
- `region`: Region where interview took place
- `polviews`: Respondent's answer to the following prompt: 

    "We hear a lot of talk these days about liberals and conservatives. I’m going to show you a seven-point scale on which the political views that people might hold are arranged from extremely liberal - point 1 - to extremely conservative - point 7. Where would you place yourself on this scale?"

Use the code below to read in the data.

```{r eval = F}
gss <- read_csv("data/gss2016.csv", 
  na = c("", "Don't know", "No answer", #assign NA if no answer 
         "Not applicable"),  
         guess_max = 2867) %>%
  select(natmass, age, sex, sei10, region, polviews) %>%
  drop_na()
```

The argument `guess_max = 2867` tells the `read_csv` function to use all of the observations in a column to determine its data type. Without this argument, only the first 1,000 observations would be used to make this determination. This becomes important for a variable like `age`; though `age` is coded as numeric data for most of the observations, there are some in which `age` is coded as `"89 or older"`. Without the `guess_max` argument, you will get warnings when loading the data.

Note also that only the variables of interest will be loaded, not the entire dataset. This will make for faster computation and knitting as you work on the lab. 

# Exercises

**Show all relevant code and output to support your responses even if you use inline code to write your narrative.**

## Part I: Exploratory data analysis

1. The goal of the analysis is to understand the factors that are associated with a person being satisfied with the current spending on mass transportation. Create a new variable that is equal to "1" if a person said spending on mass transportation is about right and "0" otherwise.

2. Recode `polviews` so it is a factor variable type with levels that are in an order that is consistent with question on the survey. *Note how the categories are spelled in the data.*

    Make a plot of the distribution of `polviews`. Which political view occurs most frequently in this data set?

3. Make a plot displaying the relationship between satisfaction with mass transportation spending  and `polviews`. Use the plot to describe the relationship between a person's political views and whether they are satisfied with spending on mass transportation. 

4. We'd like to use `age` as a quantitative variable in your model; however, it is currently a character data type because some observations are coded as `"89 or older"`. Recode `age` so that is a numeric variable. *Note: Before making the variable numeric, you will need to replace the values `"89 or older"` with a single value.*

## Part II: Logistic regression model

5. Briefly explain why we should use a logistic regression model to predict the odds a randomly selected person is satisfied with spending on mass transportation. 

6. Let's start by fitting a model using the demographic factors - `age`, `sex`, `sei10`, and `region` - to predict the odds a person is satisfied with spending on mass transportation. Make any necessary adjustments to the variables so the intercept will have a meaningful interpretation.

7. Interpret the intercept in the context of the data.

8. Consider the relationship between age and one's opinion about spending on mass transportation. Interpret the coefficient of age in terms of the **odds** of being satisfied with spending on mass transportation.

9. Now let's see whether a person's political views has a significant impact on their odds of being satisfied with spending on mass transportation, after accounting for the demographic factors. 

  Conduct a drop-in-deviance test to determine if `polviews` is a significant predictor of attitude towards spending on mass transportation. State the null and alternative hypothesis, display all relevant code and output, and state your conclusion in the context of the problem.

10. Use the model the relationship between one's political views and their attitude towards spending on mass transportation. Be sure your answer includes the interpretation of the model coefficients and associated hypothesis tests or confidence intervals used to support your response.


# Submission 

Upload the team's PDF to Gradescope. Be sure to include every team member's name in the Gradescope submission Associate the "Overall" graded section with the first page of your PDF, and mark where each answer is  to the exercises. If any answer spans multiple pages, then mark all pages.

There should only be **<u>one</u>** submission per team on Gradescope. **Be sure to include every team member's name in the Gradescope submission.**

<hr>

## Acknowledgements

The "Data" section is largely inspired by [datasciencebox.org](https://datasciencebox.org/).